---
title: "Intro to R"
author: "Jake Tolbert"
date: "April 5, 2016"
output: html_document
---

# Basics of R

1. Interactive Console

    - look at the basic console

```{r}
cyl6 <- mtcars[mtcars$cyl==6,]
cyl4 <- mtcars[mtcars$cyl==4,]

# is there a significant difference in mpg between 4 and 6 cylinder vehicles?
t.test(cyl6$mpg, cyl4$mpg)
```

    - history (just like a terminal)
    - assign and use variables (we call them objects)
    - use RStudio as an IDE instead
    

2. Running Scripts

    - save your work
    - redo analysis when your source data changes
    - avoid overwriting your objects
    

3. There's No Shame in Using Packages

    - lots of tasks are a pain in base R, munging, dates, etc
    - use libraries:
        - dplyr - general data manipulation
        - lubridate - dealing with dates
        - tidyr - tidy data and move from normalized to denormalized formats
        - ggplot2 - fancy plotting
        - magrittr - piping
        
4. Functional vs. Piping (via dplyr, magrittr)

    - R is a functional programming language, meaning everything is nested functions
    - the pipe operator %>% allows you to string commangs together linearly
        - keeps arguments togethe
        - makes code a LOT easier to read, write and think through
    
```{r echo = T}
# load the dplyr library
library(dplyr, warn.conflicts = F)

# the base R way to do it
head( mtcars[order(mtcars$hp) , c(1,2,10,4)], 10)

# using dplyr, but still way too confusing
head(arrange(select(mtcars, mpg, cyl, gear, hp), hp), 10)


# using the pipe
mtcars  %>% 
  select(mpg, cyl, gear, hp) %>% 
  arrange(hp)  %>% 
  head(10)
```
    
5. Stuff We Won't Cover

    - generating random data
    - more complex models (like random forests!)
    - general probabilty questions, density functions, etc.
        
6. "Real Life" examples with the airline data
    