---
title: "R Functions I Use"
author: "Jake Tolbert"
date: "April 5, 2016"
output: html_document
---

# R Functions I Use Regularly

## Base Functions 

`<-`  assigns values

`$`  - get a column out of a data frame, returns a vector
```{r}
# the cylinders vector of mtcars
mtcars$cyl
```

`[ ]` – get a subsection of a data frame or list
```{r}
# gets rows 2-6, columns 1:3
mtcars[2:6, 1:3] 

# returns the cyl column as a data frame
mtcars["cyl"] 
```

`[[ ]]` – get a vector that was part of a data frame or list
```{r}
# returns the cyl column as a vector
mtcars[["cyl"]] 
```
	
### Vector functions
`c()` - create a vector
```{r}
onethroughten  <- c(1:10)
onetwothree    <- c(1,2,3)
guys           <- c('Jake','RJ','Ed')
```

`summary()` - see a summary of data
```{r}
# a 5 number summary of a single vector
summary(onethroughten)

# 5 number summaries of each column of a data frame
summary(mtcars)
```

`grepl()` - return a logical vector if a vector matches a regular expression
```{r}
grepl( "setosa", iris$Species)
```

`max()`, `min()`, `mean()`, `median()`, `std()` – all do what you’d expect

`ifelse()` – Excel-style if function
```{r}
ifelse(mtcars$mpg > 30, "good mpg", "makes Green Peace sad")
```

### Plotting functions

```{r}
# plot two vectors
plot(mtcars$cyl, mtcars$mpg)

# standard boxplots
boxplot(mtcars$mpg)
boxplot(mtcars$mpg ~ mtcars$cyl)

# histograms
hist(mtcars$mpg) 
```

### Data frame functions
`data.frame()` - create a data frame from a list of vectors
```{r}
mydf  <- data.frame( 
    foo = 1:3
  , bar = c('x','y','z')
  , meh = guys
  , stringsAsFactors = F
  )
```

`str()` - see the names, classes and an excerpt of each column of a data frame

`names()` - get the column names for a data frame

`sapply()` - apply a function to each item in a list and try to return a vector
```{r}
# see the max of each column
sapply(mtcars, max)

# write your own function and return that function's value for each column
sapply(mtcars, function(x) {
  what = max(x);
  the = min(x);
  meh = what - the
  # note the lack of a return function
})
```

`lapply()` - apply a function to each item in a list and return a list
```{r}

# summarize each column
lapply(mtcars, summary)

# write a fancy function and return the value as a list
lapply(mtcars, function(x) {
        x > 10
    })
```

`unlist()` - concatenate a list into a vector

`View()`  -  see a full data frame in the RStudio viewer
```{r}
View(mtcars)
```
    

### System functions
`library(dplyr)` - load an installed package

`install.packages('dplyr')` - install a package

`if()` - flow control 
	
`exists()` - check if an object exists
```{r}
# check to make sure foo exists before you bother reading it in
if( !exists("foo") ) { read.csv("foo.csv") }
```

`rm()` - remove an object
```{r}
# delete foo
rm(foo)

# clear out the working environment
rm(list = ls() )
```

`read.csv()` - read a csv file from disk into a data frame
```{r}
# read in foo.csv
foo  <- read.csv("foo.csv", stringsAsFactors = F)

# read a tab separated spreadsheet on the clipboard
bar  <- read.csv("clipboard", sep="\t", stringsAsFactors = F) 
```
	

`write.csv()` - write a data frame to disk
```{r}
# write the object foo as foo.csv 
write.csv(foo, file= "foo.csv", row.names = F)

# write the object bar to the clipboard to paste into Excel 
write.csv(bar, file= "clipboard", sep = "\t", row.names = F)
```

## Dplyr verbs

see [(https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html)]

### Base verbs
`filter()` - filter a data frame based on criteria
```{r}
# load dplry
library(dplyr)
# make mtcars print nicely
mtcars  <- tbl_df(mtcars)


mtcars %>% filter(cyl >4)
```

`mutate()` - add new columns to a data frame
```{r}
mtcars %>% mutate( hasbigengine = cyl >= 4)
```

`select()` - choose and reorder columns in a data frame
```{r}
mtcars %>% select( cyl, drat)
```

`arrange()` - sort a data frame based on criteria in the data frame
```{r}
mtcars %>% arrange( cyl)
```

`group_by()` - set a column to aggregate on 
```{r}
foo  <- mtcars %>% group_by( cyl )
```

`summarize()` - calculate new aggregated columns on a grouped data frame
```{r}
foo %>% summarize( avg.mpg = mean(mpg) , n.in.group = n() )
```



### SQL style verbs
- `left_join()`
- `inner_join()`
- `right_join()`
- `full_join()`
- `semi_join()`
- `anti_join()`
