---
title: "Using R vs. QlikView vs. Excel"
output: html_notebook
---

Excel
===================================================

### Advantages

- You can manipulate the data itself easily
- Data is immediately visible
- Visualization tools are quick and easy

### Disadvantages 

- Any sort of formulas on large data sets are SLOW
- Not reproducible/updatable at all
- Hard to do complex visualizations
- Doing complex statisical analysis is HARD


 Qlikview
===================================================

### Adantages 

- updatable
- visualization tools are really fast
- very flexible
- fast with big data sets

### Disadvantages

- can be really messy/you've got to do all the UI work
- not reproducible - have to remember what filter you used
- can be fiddly
- statistical work is HARD
- complex visiualization are nearly impossible
- NULL value handling is WEIRD
- nonstandard formula
- sharing data with non-QV users is a mess


R
===================================================

### Advantages

- very reproducible/updatable
- building reports is simple with R Markdown
- simple to jump from data manipulation to statisical analysis
- cutting edge statistical methods are a few keystrokes away
- complex plotting is possible
- you can do anything (incl. file manipulation)
- fast with big data sets

### Disadvantages

- it's a programming language, so you have to learn it
- it's crusty if you stick with base R (this is being solved by the tidyverse)
- interactive plotting isn't really there yet
- sharing data with non-R users can be a little fiddly



Best Tool for the Job
==============================

### Pulling a mailing list from a database

- use SQL or R
  - the data changes daily, so you want to be able to save your criteria and run it against live data
  - Excel would be a nightmare
  - QV you could save a bookmark, I guess?
    - but can you do complex OR filtering in QV?
  - R allows you do that simply:
  
```{r}
mymailinglist  <- mydatabasetable %>% 
  filter(
      # do they have a mailing address
      !is.na(mailing_address) 
      
      # they haven't opted out
    , (is.na(mail_opt_out_code) | mail_opt_out_code == 'N')
    
      # is identified as target or is a lawyer in Decatur
    , (is_identified_target | (city == 'Decatur' & job_title == 'Lawyer' ) ) 
    
  ) %>% 
  select(id, name, mailing_address, city, st, zip) %>% 
  collect()
```


### Statistical Analysis 

- use R
   - runni
   
### Predictive Modeling
- use R
```{r}

summary(
t.test(
cor(


```

```{r}

lm_model     <- 
glm_model    <- 
rf_model     <- 
caret_model  <- 


```
   
### Build Executive Dashboard  

- use QlikView (or Shiny in R)

### Generate a Weekly Report with plots and tables

- use R -> Excel or R Markdown




  