---
title: "Common Tasks"
author: "Jake Tolbert"
date: "October 25, 2017"
output: html_document
---

Complex Mailing Lists
==============================

Use R (or SQL or SAS or some such). 

- Data can change underneath, but criteria can be saved.
- QV has no way to do complex filtering eg:
    - T&D majors around Decatur OR $1000 lifetime gifts to T&D
    
```{r echo=TRUE, eval = FALSE}

# imagine I'm connected to a database with databases:
#   - majors_tbl with pidm, major
#   - regions_tbl with pidm, region
#   - donors_tbl with pidm, td_gifts, athletics_gifts, uc_gifts, scholarship_gifts
#   - demographic_tbl with pidm, mailing address, and more ...

 
# get a list of all Theatre & Dance majors
tdmajors  <- majors_tbl %>% 
  filter( major == 'TD'  ) 

# get a list of everybody around Decatur
arounddecatur  <- regions_tbl %>% 
  filter( region == 'DECATUR' )

# get T&D majors around Decatur
td_around_decatur  <- tdmajors %>% 
  semi_join(arounddecatur, by = 'pidm') %>% 
  distinct(pidm)


# get donors who've given $1000+ to Theatre & Dance building
td_donors  <- donors_tbl %>% 
  filter(td_gifts  >= 1000 ) %>% 
  distinct(pidm)


#combine both lists then add mailing data
my_mailing  <- bind_rows( td_around_decatur, td_donors) %>% 
  left_join( 
    demographic_tbl %>% 
        select(pidm, name, address, city, st, zip)
      , by = 'pidm', copy = T
  ) %>% 
  collect()
```


Exploratory Analysis
==============================

When you're looking for relationships or patterns in your data, QV can be pretty good.

Load the data in, set up a bunch of list boxes and charts and click around, looking for insights. It's simple and easy and the results are immediate.



Statisical Analysis
==============================

Use R for complex statistical analysis.

```{r message=FALSE, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(
    cache = T
  , highlight = T
)


library(tidyverse)
library(lubridate)
library(magrittr)
library(randomForest)
library(caret)
library(broom)

airlinedata  <- read_csv('data/On_Time_On_Time_Performance_2017_1.csv') %>% 
  mutate( 
      FlightDate = ymd(FlightDate) 
    , Origin = factor(Origin)
    , UniqueCarrier = factor(UniqueCarrier)
  )



set.seed(500)
mydata  <- airlinedata %>% sample_n(5000) %>% 
  replace_na(list( ArrDelayMinutes = 0, ArrDelay = 0))
  
```

### Quick Summary Data
```{r}
summary(airlinedata$DepDelay)
hist(airlinedata$DepDelay)
boxplot(airlinedata$DepDelay ~ airlinedata$UniqueCarrier, ylim = c(-20,50))
```

### More Complex Analytical Tasks

```{r highlight = T}
cor.test(as.numeric(airlinedata$UniqueCarrier), airlinedata$DepDelay)

t.test(
    airlinedata %>% filter(Origin == 'ORD') %>% extract2('DepDelay')
  , airlinedata %>% filter(Origin == 'CLT') %>% extract2('DepDelay')
)
```



### Creating a Report for Others

1. Build with R Markdown - integrate Markdown + R into generated html/pdf.
2. Use QV for dashboards, etc. You can use R (using Shiny and Shiny Dashboards), but QV is designed for this sort of thing. 



### Building Statistical Models

```{r echo=T, message=FALSE, warning=FALSE}
lm_model     <- lm(ArrDelayMinutes ~ DayOfWeek + AirTime, data = mydata  )

rf_model     <- randomForest(
                      x = mydata %>% select(DayOfWeek, UniqueCarrier)
                    , y = mydata$ArrDelayMinutes
                )

caret_model  <- train(
                      x = mydata %>% select(DayOfWeek, UniqueCarrier )
                    , y = mydata$ArrDelayMinutes  
                    , trControl = trainControl(number = 3)
                )
```

```{r}

tidy(lm_model)
rf_model
caret_model

```

Once you've built your models, you can predict new cases.

```{r}
newdata  <- data_frame(
    DayOfWeek      = c(1L,2L,3L,4L,5L)
  , UniqueCarrier  = factor(c("AA", "AS", "EV", "WN", "OO"), levels = levels(mydata$UniqueCarrier))
)


predictions  <- predict(caret_model, newdata)

newdata$predictedDelay  <- predictions
newdata
```



